Corporate Security Solutions for BYOD.\\ A Novel User-Centric and Self-Adaptive System

### 2015-04-13 MUSES_vs_BYOD_solutions.tex
Changing intro to tools review

One of the tasks of a Chief Security Officer (CSO) is to elaborate an efficient set of Information Security Policies (ISPs), for controlling a certain, and already known, structure \cite{Opp_Security11}. This means that, until the appearance of BYOD, enterprises used to follow static security policies devoted to control an environment where the company assets and the devices were purchased and maintained by the company. Now that corporate networks are becoming dynamic for being adapted to this BYOD philosophy, there is an additional risk because the devices that the employees use are not always company owned. This means that the CSOs have to find the balance between having a fast response to any user action that might cause harm (in terms of money loss because of a security incident), and trying to avoid monitoring the users in a way that is against privacy. A needed security policy, or in this case, an ISP should deal with the way of protecting a certain organization's information against a security breach. Though there are standards, such as the ISO27002 or the Security Forum's Standard of Good Practice\footnote{\url{https://www.securityforum.org}}, and many guidelines \cite{SecPol09}, an ISP should be adapted depending on the characteristics of the community/organisation that they are built for. In this sense, MUSES stands up for a self-adaptation of the security policies of a company.

On the other hand, employee owned devices like smartphones, can be both used for personal matters at work, and for continue working outside of the workplace. As the name says, smartphones are more than simple mobile phones, and people who use them in their works have the possibility of maintaining a good balance between work and private life. For this reason, the risk of uncontrolled devices accessing to corporate assets in unsafe conditions, due to the number of new risks which are linked to smartphones \cite{gangula2013survey}, is bigger.

@@

Thus now, more things should be considered than the usual ones when designing a company's network architecture. In Figure \ref{fig:proposed_diagram} there is a proposal which can be used for the beginning of the study of solutions that may secure such a dynamic environment. It includes the possibility of having employee-owned mobile (smartphones and tablets) and portable (laptops) devices, and also the opportunity that the employees have of connecting these devices either from inside or outside the company premises. Moreover, company's information assets are constantly accessed under these conditions, considering that an information asset means every \textit{piece of information} that has a \textit{value} (cost depending on the risk of being lost or leaked) for the company. It can be referred to files with sensitive information to certain mails, or even to company applications.

\begin{figure}[ht]
	\begin{center}
		\includegraphics[scale=0.4]{img/proposed_diagram.eps}
		\caption{Architecture approach of an Enterprise Network assuming that the Company has adopted the BYOD philosophy.}
	\label{fig:proposed_diagram}
	\end{center}
\end{figure}

@@

Now that BYOD philosophy is becoming a tendency, a number of tools appeared by now, and others are under development but expected to be released soon. They were specifically designed for Chief Security Officers (CSOs) and Chief Information Security Officers (CISOs) to secure, monitor, and act over smartphones and other personal mobile or portable devices, when they become too risky. Some of those tools have influenced the development of the MUSES system  itself (Section \ref{sec:muses}). The present section introduces the most relevant features of these products, as they can be considered related to the MUSES objectives.

@@

One of the first companies who supported the BYOD model was IBM in 2011 \cite{IBM_tool}, as they recognized the increase of employees who brought their personal smartphones or tablets into the workplace. To help organizations \cite{ibm11} embracing both company and employee owned mobile devices (as said, this practise is part of the BYOD model) in a security-rich environment, IBM developed a mobile device security management solution. For IBM, a mobile security strategy should focus on several key areas.

@@

\subsubsection{Azzurri - Icon Mobilise \cite{AzurriMDM}}
It is a cloud-based service to manage enterprise devices from the BYOD point of view (protecting sensitive corporate data and also letting the users enjoy them privately), by outsourcing the management of those devices to Azzurri. 
\begin{itemize}
\item It offers a centrally deployment, administration and securing of all mobile devices regardless of their OS.
\item It manages and enforces security policies `over the air' for both corporate and employee owned mobile devices.
\item It does it by enforcing passwords and providing mechanisms for remote lock.
\item It offers the ability to wipe/selective wipe corporate data email/contacts on the device if lost or stolen.
\end{itemize}

>>>>

One of the tasks of a Chief Security Officer (CSO) is to elaborate an efficient set of Information Security Policies (ISPs), for controlling a certain, and already known, structure \cite{Opp_Security11}. This means that, until the appearance of BYOD, enterprises used to follow static security policies devoted to control an environment where the company assets and the devices were purchased and maintained by the company. Now that corporate networks are becoming dynamic for being adapted to this BYOD philosophy, there is an additional risk because the devices that the employees use are not always company-owned. This means that the CSOs have to find the balance between having a fast response to any user action that might cause harm (in terms of money loss because of a security incident), and trying to avoid monitoring the users in a way that is against privacy. A needed security policy, or in this case, an ISP should deal with the way of protecting a certain organization's information against a security breach. Though there are standards, such as the ISO27002 or the Security Forum's Standard of Good Practice\footnote{\url{https://www.securityforum.org}}, and many guidelines \cite{SecPol09}, an ISP should be adapted depending on the characteristics of the community/organisation that they are built for. In this sense, MUSES stands up for a self-adaptation of the security policies of a company.

On the other hand, employee-owned devices like smartphones, can be both used for personal matters at work, and for continue working outside of the workplace. As the name says, smartphones are more than simple mobile phones, and people who use them in their works have the possibility of maintaining a good balance between work and private life. For this reason, the risk of uncontrolled devices accessing to corporate assets in unsafe conditions, due to the number of new risks which are linked to smartphones \cite{gangula2013survey}, is bigger.

@@

Thus now, more risky situations should be considered when designing a company's network architecture. In Figure \ref{fig:proposed_diagram} there is a proposal which can be used for the beginning of the study of solutions that may secure such a dynamic environment. It includes the possibility of having employee-owned mobile (smartphones and tablets) and portable (laptops) devices, and also the opportunity that the employees have of connecting these devices either from inside or outside the company premises. Moreover, company's information assets are constantly accessed under these conditions, considering that an information asset means every \textit{piece of information} that has a \textit{value} (cost depending on the risk of being lost or leaked) for the company. It can be referred to files with sensitive information, as well as confidential mails, or even to company applications.

\begin{figure}[ht]
	\begin{center}
		\includegraphics[scale=0.4]{img/proposed_diagram.eps}
		\caption{Architecture approach of an enterprise network, assuming that it has adopted the BYOD philosophy.}
	\label{fig:proposed_diagram}
	\end{center}
\end{figure}

@@

Many tools for companies, as well as for devices, which have adopted the BYOD have been released in the past four years. This way, and
more focused on the enterprise, offer the CSO ways to control the devices which enter in the system, requiring users to employ strong passwords, for instance, and also to protect the employees data by means of data encryption and data protection by having strong and secure passwords. Other tools for managing a BYOD situation add to their features guidelines for the CSOs to develop good ISPs. On the other hand, many solutions have been presented which are more focused on the device side, although they implement also the server side. Some of those tools have influenced the development of the MUSES system  itself (Section \ref{sec:muses}), which can complement them, as adds other features that go beyond the state of the art. The present section introduces the most relevant features of these products, as they can be considered related to the MUSES objectives.

@@

One of the first companies who supported the BYOD model was IBM in 2011 \cite{IBM_tool}, as they recognized the increase of employees who brought their personal smartphones or tablets into the workplace. To help organizations \cite{ibm11} embracing both company and employee-owned mobile devices (as said, this practise is part of the BYOD model) in a security-rich environment, IBM developed a mobile device security management solution. For IBM, a mobile security strategy should focus on several key areas.

@@

\subsubsection{Azzurri - Icon Mobilise \cite{AzurriMDM}}
It is a cloud-based service to manage enterprise devices from the BYOD point of view (protecting sensitive corporate data and also letting the users enjoy them privately), by outsourcing the management of those devices to Azzurri. 
\begin{itemize}
\item It offers a centrally deployment, administration and securing of all mobile devices regardless of their OS.
\item It manages and enforces security policies `over the air' for both corporate and employee-owned mobile devices.
\item It does it by enforcing passwords and providing mechanisms for remote lock.
\item It offers the ability to wipe/selective wipe corporate data email/contacts on the device if lost or stolen.
\end{itemize}
